# Coursera Reproducible Research Week 2 Project 1

## Code for reading in the dataset and/or processing the data
```{r, echo=TRUE}
stepsdata <- read.csv("C:\\Users\\sanjiv\\Desktop\\activity.csv", header = TRUE, sep = ",")
head(stepsdata)
```

## Histogram of the total number of steps taken each day

```{r}
library("dplyr")
newdata <- stepsdata %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps = sum(steps)) %>% na.omit()
hist(newdata$tsteps, xlab = "Total steps per day", main = "daily total steps", breaks = 53)
```

## Mean and median number of steps taken each day

```{r}
mean(newdata$tsteps)
median(newdata$tsteps)
```

## Time series plot of the average number of steps taken

```{r}
library(ggplot2)
data_interval <- stepsdata %>% select(interval, steps) %>%na.omit() %>% group_by(interval) %>% summarize(tsteps = mean(steps))
ggplot(data_interval, aes(interval, tsteps)) + geom_line()
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r}
data_interval[which(data_interval$tsteps== max(data_interval$tsteps)),]
```

## Code to describe and show a strategy for imputing missing data

```{r}
missingval <- sum(is.na(stepsdata))
missingval
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepsdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
newdatabydate <- meandata %>% select(date, steps) %>% group_by(date) %>% summarize(totalsteps = sum(steps))
hist(newdatabydate$totalsteps, xlab = "total steps", ylab = "frequency", main = "Total daily steps", breaks = 61)
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )
meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")
 
ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
     facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
     ggtitle("Comparison of Average Number of Steps in Each Interval")
```

##Fin. 


